Supporting information: Tail associations and their impact on extinction risk

### 2019-10-29 SuppMat.Rmd
edited notations

dt2<-tibble("Density dependence $(\\lambda_t^{i})$ with noise $(\\epsilon_t^{i})$" = c("$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} \\exp[r (1-P_t^{i}/K)$]",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/(1+aP_t^{i})^b$",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/[1+(aP_t^{i})^b]$",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} r/[1+\\exp \\{-a(1-P_t^{i}/b)\\}]$",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} [1+r(1-P_t^{i}/K)]$",
                          "$\\exp(\\epsilon_t^{i}) \\hspace{0.1 cm} [1+r(K-P_t^{i})\\{1-\\exp(-sP_t^{i})\\}]$"
))

@@

dt4<-tibble("$P_e^{i}$" = c("K",
                        "$(r^{1/b}-1)/a$",
                        "$[(r-1)^{1/b}]/a$",
                        "$[b\\{a+ln(r-1)\\}]/a$",
                        "K",
                        "K"
))

dt5<-tibble("$r_c^{i}$" = c(rc_ricker,
                        rc_hassell,
                        rc_msmith,
                        rc_pennycuick,
                        rc_verhulst,
                        rc_abrewer
))

dt6<-tibble("$r_{bf}^{i}$" = c(r_bf_ricker,
                           r_bf_hassell,
                           r_bf_msmith,
                           r_bf_pennycuick,
                           r_bf_verhulst,
                           r_bf_abrewer
))

@@

<!--FIG: Ricker model, local dispersal-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
standard deviation $\sigma$=1 was used to generate extreme extreme tail dependent noises for 10000 simulations. \label{fig_ricker_risk_local_disp_scl_1}}
\end{center}
\end{figure}

<!--FIG : Hassell model, local dispersal-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
\includegraphics[width=15 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_hassell_risk_local_disp_scl_1}}
\end{center}
\end{figure}

<!--FIG: Maynard smith model, local dispersal-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
\includegraphics[width=15 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Maynard smith metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
standard deviation $\sigma$=0.8 was used to generate extreme tail dependent noises for 10000 simulations.. \label{fig_msmith_risk_local_disp_scl_0.8}}
\end{center}
\end{figure}

<!--FIG: Pennycuick model, local dispersal-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
\includegraphics[width=15 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
\end{center}
\end{figure}

<!--FIG: Verhulst model, local dispersal-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
\end{center}
\end{figure}

<!--FIG: Austin-Brewer model, local dispersal-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
\caption[Effect of extreme tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Effect of extreme tail-dependent environmental noises on extinction risk (calculated after 25 years)
for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
\end{center}
\end{figure}

@@

The moderate tail association was modeled by using the 
concept of "copula", a statistical tool to model 
dependencies. Appendix \ref{SM-intro_copula} summarized
a brief introduction of "copulas" and their relation to tail associations, symmetries 
with examples of available copula families 
from `R` packages (`copula` and `VineCopula`). We used bivariate Clayton 
copula to model moderate left-tail and Survival Clayton copula to model 
moderate right-tail association in the vital rates. Both copulae had the same Kendall 
correlation ($\rho =$ `r srho_common`) but opposite association in their respective tails. 
We generated a $N \times 2$ matrix, M, for each of the left (Clayton) and 
right (Survival Clayton) tail associated bivariate copulas (see `BiCopSim` function from
`VineCopula` package ), where $N$ is number of years ($=25$) multiplied 
by number of simulations ($=10000$) we used. 
We chose parameters for each copula family using 
`iRho` function from `copula` package so that both copula had
same Kendall's correlation ($\rho =$ `r srho_common`). Entries within each column of
matrix $M$, then became independent of each other but 
entries between columns had a specific tail association
with positive Kendall's correlation.

@@

For moderately left-tail associated 
noises we drew points from a $n$ dimensional 
Clayton copula with same Kendall correlation ($\rho =$ `r srho_common`) between
any two out of $n$ patches and made 
the marginals normal with zero mean and standard deviation $\sigma$ 
(see the specified value $\sigma$ for each model in 3$^{rd}$ 
column of Table \ref{tab_model_summary}). For moderately 
right-tail associated noises we flipped ($180^{\circ}$ rotation) the points drawn 
from Clayton copula so that all the statistical properties remained 
same except for the direction of tail association. We generated
extreme left- or right-tail associated noises via algorithm stated in 
Appendix \ref{SM-extr_tail_dep}.

@@

First, we will briefly discuss about the stability analysis for a difference 
equation which is frequently used in population ecology model. Let's say, $P_{t+1} = F(P_t,r)$
represents the model dynamics of a single population $P$ which progress recursively via 
a nonlinear density-dependence function $F$ with discrete time variable $t$. $r$ is the model parameter 
which plays major role in assessing the model's dynamical stability. Fixed-point analysis can
be used to get information about this type of nonlinear difference equation. The idea is to
employ the concept that the system at fixed point (or equilibrium $P_e$) does not change with time 
(i.e. $P_{t+1} = P_t$). Given a tiny perturbation from its equilibrium $P_e$, if the system 
returns to its original state, then $P_e$ is called to be a stable equilibrium, otherwise unstable.
Mathematically, from standard linear stability analysis we can say an equilibrium is locally 
asymptotically stable if $| {F}' | < 1$, where ${F}' = \left( \frac{\mathrm{d}F }{\mathrm{d} P} \right) _{P_t=P_e}$.
If the approach towards its equilibrium is monotonic (when $0<{F}'<1$), the dynamics is said to be
in under-compensatory regime. An over-compensatory dynamics means the approach towards its stable equlibrium
is oscillatory ($-1<{F}'<0$). Following we will describe stability condition of the model dynamics considered for each 
patch of metapopulation models in the maintext (see Table \ref{MT-tab_model_summary}).

**Ricker model** [@ricker1954stock; @may1976bifurcations], originally introduced in the context of modelling
stock and recruitment of salmon populations, is a classic 
discrete population model 
and frequently used in fisheries management problem. The model dynamics
is represented as 
$P_{t+1}=F(P_t,r,K)=P_t \exp[r(1-P_t/K)]$, where $P$ is the population density, $r>0$ is the intrinsic 
growth rate and $K>0$ is the carrying capacity. There are two equilibrium states ($P_e = 0, K$).
The equilibrium at ($P_e = 0$) is unstable as ${F}' = \exp(r) > 1$ for a positive $r$ and considered 
as the extinction equilibrium 
in population ecology context. Depending on parameter $r$, 
the other equilibrium $P_e = K$ can be stable or unstable as ${F}' = (1-r)$. Following generalized
concept, the model can show monotonic approach to its stable equilibrium $P_e = K$ if $0<r<1$ 
or can approach $P_e = K$ in an oscillating fashion if $1<r<2$. When 
$r>2$ (i.e., ${F}'<-1$), system shows sustained oscillation and the equilibrium becomes unstable. We called $r=1$ 
as $r_c$, the critical point where system goes from under- to over-compensatory regime and 
$r=2$ as $r_{bf}$ where bifurcation (transition from stable to unstable equilibrium) occurs.

**Hassell model** [@hassell1975density; @may1976bifurcations], motivated for modelling insect populations,
can be represented as $P_{t+1} = F(P_t,r,a,b)=rP_t/(1+aP_t)^b$, where $r>0$ is the intrinsic growth rate 
of population density $(P)$. $a$ and $b$ are the other positive parameters. If $b=1$, it reduces to
Beverton-Holt model [@beverton2012dynamics] used in fisheries. In general, if $b \sim 1$,
then we say that there is contest and $F$ will saturate eventually at a finite
level for large populations, which indicates that the population stabilizes by
rejecting too many newborns. For scramble, $b$ is large and $F$ tends to
zero for large populations which indicates that the resources are over-utilized
leading to eventual extinction. Like Ricker, Hassel model too has two equilibrim:
$P_e =0$ (trivial extinction equilibrium) and $P_e = P^\star =(r^{1/b}-1)/a$. 
For $P_e = P^\star$, model remains in under-compensatory regime if $1<r<[b/(b-1)]^b$ 
and in over-compensatory regime if $[b/(b-1)]^b<r<[b/(b-2)]^b$ with $r_c = [b/(b-1)]^b$.
Hopf bifurcation occurs at $r_{bf} = [b/(b-2)]^b$, so if $r>r_{bf}$, $P_e=P^\star$
becomes unstable equilibrium.

**Maynard Smith model** [@cohen1995unexpected; @maynard1978models; @may1976bifurcations] follows the dynamics 
$P_{t+1}=F(P_t,r,a,b)=rP_t/[1+(aP_t)^b]$. $r$ is the growth rate, $a, b$ are the parameters with positive values.
Equilibriums are: $P_e=0$(as expected extinction equilibrium) and $P_e = P^\star=\frac{1}{a}(r-1)^{1/b}$.
For a positive equilibrium value of $P^\star$, $r$ must be greater than $1$. This $P^\star$ will be stable
in under-compensatory regime for $1<r<r_c(=b/(b-1))$ and in over-compensatory regime for $r_c<r<b/(b-2)$.
$P^\star$ will become unstable when $r>r_{bf}(=b/(b-2))$.

**Pennycuick model** [@pennycuick1968computer; @may1976bifurcations] was 
initially developed
to follow change in age-structured total population
level which assumed
to have a constant sex 
ratio and therefore change in female population was 
expected through the change 
in total population density. 
A two population variant could also be possible
in prey-predator perspective as indicated in [@pennycuick1968computer].
In general, model dynamics is expressed as $P_{t+1}=F(P_t,r,a,b)=rP_t/[1+\exp \left\{-a(1-P_t/b)\right\}]$, where 
$r$ is the intrinsic growth rate and $a>0,b>0$ are the model parameters. Model
equilibriums are 
$P_e=0$ (extinction equilibrium) and $P_e=P^\star=\frac{b}{a}[a+\ln{(r-1)}]$. For 
a finite equilibrium population 
$r$ must be greater than $r_{min}$, where 
$r_{min} = 1+\exp(-a)$. The critical transition point $r_c$ (from under- to 
over-compensatory dynamics) and Hopf-bifurcation point $r_{bf}$ can be obtained
by numerical solution of ${F}'=0$ and ${F}'=-1$ at $P_e=P^{\star}$, respectively.

**Verhulst model** [@cohen1995unexpected; @may1976bifurcations] is a discrete time difference 
representation of logistic growth of non-human biological population and can be written 
as $P_{t+1} = F(P_t,r,K) =P_t[1+r(1-P_t/K)]$, wirh $r>0$ and $K>0$ being growth rate and carrying capacity, respectively. 
Model equilibrium $P_e=0$ is always unstable whereas $P_e=K$ is stable below Hopf-bifurcation 
point $r_{bf}(=2)$. Model at stable equilibrium ($P_e=K$) lies in the under-compensatory regime when $0<r<r_c$
and in over-compensatory regime when $rc<r<r_{bf}$ with $r_c$ being equal to 1.

**Austin-Brewer model** [@austin1971world; @cohen1995unexpected], unlike from the first
five models we considered,
is used to model growth of human population. It follows the
discrete dynamics : $P_{t+1} = F(P_t,r,K,s)= P_t[1+r(K-P_t) \left\{ 1-\exp (-sP_t)\right\}]$, where
$r, K, s$ are positive, representing growth rate, carrying capacity 
and model parameter, respectively.
Usual extinction equilibrium is $P_e=0$. The other
finite, positive equilibrium is $P_e=P^\star=1-rK\left\{ 1-\exp (-sK)\right\}$. The equilibrium $P^\star$
is stable when $r<r_{bf}$ and unstable beyond the bifurcatio point $r_{bf} = \frac{2}{K[1-\exp (-sK)]}$.
After perturbation, model can return back monotonically to its stable equilibrium if parameter $r$ lies in the 
under-compensatory regime ($0<r<r_c$), where $r_c = \frac{1}{K[1-\exp (-sK)]}$. In the 
over-compensatory regime ($r_c<r<r_{bf}$), model approaches to stable equilibrium $P^\star$
in oscillatory fashion.

>>>>

dt2<-tibble("Density dependence $(\\lambda_i(t))$ with noise $(\\epsilon_i(t))$" = c("$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} \\exp[r (1-P_i(t)/K)$]",
                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} r/(1+aP_i(t))^b$",
                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} r/[1+(aP_i(t))^b]$",
                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} r/[1+\\exp \\{-a(1-P_i(t)/b)\\}]$",
                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} [1+r(1-P_i(t)/K)]$",
                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} [1+r(K-P_i(t))\\{1-\\exp(-sP_i(t))\\}]$"
))

@@

dt4<-tibble("$P_{e}$" = c("K",
                        "$(r^{1/b}-1)/a$",
                        "$[(r-1)^{1/b}]/a$",
                        "$[b\\{a+ln(r-1)\\}]/a$",
                        "K",
                        "K"
))

dt5<-tibble("$r_{c}$" = c(rc_ricker,
                        rc_hassell,
                        rc_msmith,
                        rc_pennycuick,
                        rc_verhulst,
                        rc_abrewer
))

dt6<-tibble("$r_{bf}$" = c(r_bf_ricker,
                           r_bf_hassell,
                           r_bf_msmith,
                           r_bf_pennycuick,
                           r_bf_verhulst,
                           r_bf_abrewer
))

@@

<!--FIG: Ricker model, local dispersal-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
\caption[Effect of tail association in environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years)
for $5$ metapopulation patches which followed Ricker dynamics with local dispersal; 
standard deviation $\sigma$=1 was used to generate extreme extreme tail dependent noises for 10000 simulations. \label{fig_ricker_risk_local_disp_scl_1}}
\end{center}
\end{figure}

<!--FIG : Hassell model, local dispersal-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
\includegraphics[width=15 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
\caption[Effect of tail association in environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years)
for $5$ metapopulation patches which followed Hassell dynamics with local dispersal; 
standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_hassell_risk_local_disp_scl_1}}
\end{center}
\end{figure}

<!--FIG: Maynard smith model, local dispersal-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
\includegraphics[width=15 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
\caption[Effect of tail association in environmental noises on extinction risk for Maynard smith metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years) for $5$ metapopulation patches which followed Maynard Smith dynamics with local dispersal; 
standard deviation $\sigma$=0.8 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_msmith_risk_local_disp_scl_0.8}}
\end{center}
\end{figure}

<!--FIG: Pennycuick model, local dispersal-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
\includegraphics[width=15 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
\caption[Effect of tail association in environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years)
for $5$ metapopulation patches which followed Pennycuick dynamics with local dispersal; 
standard deviation $\sigma$=1 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1}}
\end{center}
\end{figure}

<!--FIG: Verhulst model, local dispersal-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
\caption[Effect of tail association in environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years)
for $5$ metapopulation patches which followed Verhulst dynamics with local dispersal; 
standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6}}
\end{center}
\end{figure}

<!--FIG: Austin-Brewer model, local dispersal-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp_extreme_and_moderate_noise.pdf}\\
\caption[Effect of tail association in environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Effect of tail association in environmental noises on extinction risk (calculated after 25 years) for $5$ metapopulation patches which followed Austin-Brewer dynamics with local dispersal; 
standard deviation $\sigma$=0.6 was used to generate extreme tail dependent noises for 10000 simulations. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6}}
\end{center}
\end{figure}

@@

The moderate tail association was modeled by using the 
concept of "copula", a statistical tool to model 
dependencies. Appendix \ref{SM-intro_copula} summarized
a brief introduction of "copulas" and their relation to tail associations, symmetries 
with examples of available copula families 
from `R` packages (`copula` and `VineCopula`). We used bivariate Clayton 
copula to model moderate left-tail and Survival Clayton copula to model 
moderate right-tail association in the vital rates. Both copulae had the same Spearman 
correlation ($\rho =$ `r srho_common`) but opposite association in their respective tails. 
We generated a $N \times 2$ matrix, M, for each of the left (Clayton) and 
right (Survival Clayton) tail associated bivariate copulas (see `BiCopSim` function from
`VineCopula` package ), where $N$ is number of years ($=25$) multiplied 
by number of simulations ($=10000$) we used. 
We chose parameters for each copula family using 
`iRho` function from `copula` package so that both copula had
same Spearman correlation ($\rho =$ `r srho_common`). Entries within each column of
matrix $M$, then became independent of each other but 
entries between columns had a specific tail association
with positive Spearman correlation.

@@

For moderately left-tail associated 
noises we drew points from a $n$ dimensional 
Clayton copula with same Spearman correlation ($\rho =$ `r srho_common`) between
any two out of $n$ patches and made 
the marginals normal with zero mean and standard deviation $\sigma$ 
(see the specified value $\sigma$ for each model in 3$^{rd}$ 
column of Table \ref{tab_model_summary}). For moderately 
right-tail associated noises we flipped ($180^{\circ}$ rotation) the points drawn 
from Clayton copula so that all the statistical properties remained 
same except for the direction of tail association. We generated
extreme left- or right-tail associated noises via algorithm stated in 
Appendix \ref{SM-extr_tail_dep}.

@@

First, we will briefly discuss about the stability analysis for a difference 
equation which is frequently used in population ecology model. Let's say, $P(t+1) = F(P(t),r)$
represents the model dynamics of a single population $P$ which progress recursively via 
a nonlinear density-dependence function $F$ with discrete time variable $t$. $r$ is the model parameter 
which plays major role in assessing the model's dynamical stability. Fixed-point analysis can
be used to get information about this type of nonlinear difference equation. The idea is to
employ the concept that the system at fixed point (or equilibrium $P_e$) does not change with time 
(i.e. $P(t+1) = P(t)$). Given a tiny perturbation from its equilibrium $P_e$, if the system 
returns to its initial state, then $P_e$ is called to be a stable equilibrium, otherwise unstable.
Mathematically, from standard linear stability analysis we can say an equilibrium is locally 
asymptotically stable if $| {F}' | < 1$, where ${F}' = \left( \frac{\mathrm{d}F }{\mathrm{d} P} \right) _{P(t)=P_e}$.
If the approach towards its equilibrium is monotonic (when $0<{F}'<1$), the dynamics is said to be
in under-compensatory regime. An over-compensatory dynamics means the approach towards its stable equlibrium
is oscillatory ($-1<{F}'<0$). Following we will describe stability condition of the model dynamics considered for each 
patch of metapopulation models in the maintext (see Table \ref{MT-tab_model_summary}).

**Ricker model** [@ricker1954stock; @may1976bifurcations], originally introduced in the context of modelling
stock and recruitment of salmon populations, is a classic 
discrete population model 
and frequently used in fisheries management problem. The model dynamics
is represented as 
$P(t+1)=F(P(t),r,K)=P(t) \exp[r(1-P(t)/K)]$, where $P$ is the population density, $r>0$ is the intrinsic 
growth rate and $K>0$ is the carrying capacity. There are two equilibrium states ($P_e = 0, K$).
The equilibrium at ($P_e = 0$) is unstable as ${F}' = \exp(r) > 1$ for a positive $r$ and considered 
as the extinction equilibrium 
in population ecology context. Depending on parameter $r$, 
the other equilibrium $P_e = K$ can be stable or unstable as ${F}' = (1-r)$. Following generalized
concept, the model can show monotonic approach to its stable equilibrium $P_e = K$ if $0<r<1$ 
or can approach $P_e = K$ in an oscillating fashion if $1<r<2$. When 
$r>2$ (i.e., ${F}'<-1$), system shows sustained oscillation and the equilibrium becomes unstable. We called $r=1$ 
as $r_c$, the critical point where system goes from under- to over-compensatory regime and 
$r=2$ as $r_{bf}$ where bifurcation (transition from stable to unstable equilibrium) occurs.

**Hassell model** [@hassell1975density; @may1976bifurcations], motivated for modelling insect populations,
can be represented as $P(t+1) = F(P(t),r,a,b)=rP(t)/(1+aP(t))^b$, where $r>0$ is the intrinsic growth rate 
of population density $(P)$. $a$ and $b$ are the other positive parameters. If $b=1$, it reduces to
Beverton-Holt model [@beverton2012dynamics] used in fisheries. In general, if $b \sim 1$,
then we say that there is contest and $F$ will saturate eventually at a finite
level for large populations, which indicates that the population stabilizes by
rejecting too many newborns. For scramble, $b$ is large and $F$ tends to
zero for large populations which indicates that the resources are over-utilized
leading to eventual extinction. Like Ricker, Hassel model too has two equilibrim:
$P_e =0$ (trivial extinction equilibrium) and $P_e = P^\star =(r^{1/b}-1)/a$. 
For $P_e = P^\star$, model remains in under-compensatory regime if $1<r<[b/(b-1)]^b$ 
and in over-compensatory regime if $[b/(b-1)]^b<r<[b/(b-2)]^b$ with $r_c = [b/(b-1)]^b$.
Hopf bifurcation occurs at $r_{bf} = [b/(b-2)]^b$, so if $r>r_{bf}$, $P_e=P^\star$
becomes unstable equilibrium.

**Maynard Smith model** [@cohen1995unexpected; @maynard1978models; @may1976bifurcations] follows the dynamics 
$P(t+1)=F(P(t),r,a,b)=rP(t)/[1+(aP(t))^b]$. $r$ is the growth rate, $a, b$ are the parameters with positive values.
Equilibriums are: $P_e=0$(as expected extinction equilibrium) and $P_e = P^\star=\frac{1}{a}(r-1)^{1/b}$.
For a positive equilibrium value of $P^\star$, $r$ must be greater than $1$. This $P^\star$ will be stable
in under-compensatory regime for $1<r<r_c(=b/(b-1))$ and in over-compensatory regime for $r_c<r<b/(b-2)$.
$P^\star$ will become unstable when $r>r_{bf}(=b/(b-2))$.

**Pennycuick model** [@pennycuick1968computer; @may1976bifurcations] was 
initially developed
to follow change in age-structured total population
level which assumed
to have a constant sex 
ratio and therefore change in female population was 
expected through the change 
in total population density. 
A two population variant could also be possible
in prey-predator perspective as indicated in [@pennycuick1968computer].
In general, model dynamics is expressed as $P(t+1)=F(P(t),r,a,b)=rP(t)/[1+\exp \left\{-a(1-P(t)/b)\right\}]$, where 
$r$ is the intrinsic growth rate and $a>0,b>0$ are the model parameters. Model
equilibriums are 
$P_e=0$ (extinction equilibrium) and $P_e=P^\star=\frac{b}{a}[a+\ln{(r-1)}]$. For 
a finite equilibrium population 
$r$ must be greater than $r_{min}$, where 
$r_{min} = 1+\exp(-a)$. The critical transition point $r_c$ (from under- to 
over-compensatory dynamics) and Hopf-bifurcation point $r_{bf}$ can be obtained
by numerical solution of ${F}'=0$ and ${F}'=-1$ at $P_e=P^{\star}$, respectively.

**Verhulst model** [@cohen1995unexpected; @may1976bifurcations] is a discrete time difference 
representation of logistic growth of non-human biological population and can be written 
as $P(t+1) = F(P(t),r,K) =P(t)[1+r(1-P(t)/K)]$, with $r>0$ and $K>0$ being growth rate and carrying capacity, respectively. 
Model equilibrium $P_e=0$ is always unstable whereas $P_e=K$ is stable below Hopf-bifurcation 
point $r_{bf}(=2)$. Model at stable equilibrium ($P_e=K$) lies in the under-compensatory regime when $0<r<r_c$
and in over-compensatory regime when $rc<r<r_{bf}$ with $r_c$ being equal to 1.

**Austin-Brewer model** [@austin1971world; @cohen1995unexpected], unlike from the first
five models we considered,
is used to model growth of human population. It follows the
discrete dynamics : $P(t+1) = F(P(t),r,K,s)= P(t)[1+r(K-P(t)) \left\{ 1-\exp (-sP(t))\right\}]$, where
$r, K, s$ are positive, representing growth rate, carrying capacity 
and model parameter, respectively.
Usual extinction equilibrium is $P_e=0$. The other
finite, positive equilibrium is $P_e=P^\star=1-rK\left\{ 1-\exp (-sK)\right\}$. The equilibrium $P^\star$
is stable when $r<r_{bf}$ and unstable beyond the bifurcatio point $r_{bf} = \frac{2}{K[1-\exp (-sK)]}$.
After perturbation, model can return back monotonically to its stable equilibrium if parameter $r$ lies in the 
under-compensatory regime ($0<r<r_c$), where $r_c = \frac{1}{K[1-\exp (-sK)]}$. In the 
over-compensatory regime ($r_c<r<r_{bf}$), model approaches to stable equilibrium $P^\star$
in oscillatory fashion.
